---
title: "reformatting-ny-data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reformatting-ny-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(precinctsopenelex)
library(dplyr)
library(tidyr)
library(stringr)
library(janitor)
library(readxl)
```

This vignette is intended to provide a walk-through of processing a set of precinct results from Saratoga County, NY, for the 2020 U.S. presidential election.  
  
This sample dataset for Saratoga is included with the package to allow for exploration, and provides a template for how the data should be structured for the functions to operate.  

```{r}
#chose state abbreviation and county name variable needed for the input file string
target_state <- "NY"
target_county <- "Saratoga"

#we'll create a name string using a package function to import the data file from Excel based on naming conventions used
infile_string <- create_infile_string(target_state, target_county)

infile_string
```

We'll use the included sample dataset - "precinctsampledata_ny" - to demonstrate that's been cleaned and ready for processing.

```{r}
head(precinctsampledata_ny)
glimpse(precinctsampledata_ny)
```

Note the precise way of formatting the column names that will allow the function to work correctly:  
* candidate columns should be written as "candidate name - party"  
* additional choices should be listed as "WriteIn" (write-in votes), "Blanks" (undervotes), and "Voids" (overvotes)

```{r}



#now let's run the processing to reformat the data into the open elections standardized format, which is tidy/long
#as opposed to wide

# Function wants:
# - dataset (or import from excel function)
# - office: text label for office (e.g. "U.S. House")
# - district: text label for district (e.g. "42"
#   note that presidential has no district, it's just "")

# *note: the import function itself can be wrapped inside, as long as the data coming in is the correct structure
# processed_prez <- process_ny_data(read_excel(filestring_import, sheet = "presidential"), 
#                                   "President", 
#                                   "")

processed_prez <- reshape_ny_data(precinctsampledata_ny,
                                  "Presidential",
                                  "")

head(processed_prez, 10)




```


